---
title: "Computer lab 2 block 2 - Version 2"
author: "Phillip Hölscher"
date: "29 7 2019"
output: 
  html_document:
    toc: True
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Assignment 1. Using GAM and GLM to examine the mortaily rates

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}
# list of all libraries
#install.packages("readxl") # it??s an xlsx file - need the readxl package for this
library(readxl)
library(ggplot2)
#install.packages("mgcv")
library(mgcv)
```


```{r}
# set working directory
#setwd("X")
```

## 1.1 Time seires plot
```{r, warning=FALSE, echo=FALSE, message=FALSE}

## 1.1 Time seires plot

data <- read_excel("Influenza.xlsx")
```


```{r, echo=FALSE}
# time series plot with dual y - axes
ggplot(data, aes(x = data$Time)) + 
  geom_line(aes(y = data$Influenza, color = "Influenza")) +
  geom_line(aes(y = data$Mortality/15, color = "Mortality")) +
  xlab("Time") + ggtitle("Mortality vs Influenza") + ylab("Influenza") + theme_bw() +
  scale_y_continuous(sec.axis = sec_axis(~.*15, name = "Mortality"))
```


## 1.2 Fit GAM model

```{r, echo=FALSE}

## 1.2 Fit GAM model

# fit GAM model
# https://www.rdocumentation.org/packages/mgcv/versions/1.8-26/topics/gam
```


```{r, echo=TRUE}
# fit GAM model
gam_fit = gam(data$Mortality ~ data$Year +
                s(data$Week, k=length(unique(data$Week)),
                  bs = "cp"), # bs = "cp" ???
              data = data,
              family = gaussian, # gaussian is defult
              method = "GCV.Cp") # method generalized cross-validation
# method : The smoothing parameter estimation method
# "GCV.Cp" to use GCV for unknown scale parameter
# bs: B-Spline Basis for Polynomial Splines

```

Probabilistic model:
$$y = Mortality$$
$$y = N(\mu, \sigma^2) $$
$$\hat{y} = \beta_0 + \beta_1Year_i + f(Week_i) + \epsilon_i$$

## 1.3 Predicted vs observed mortality + fit GAM model
```{r, echo=FALSE}

## 1.3 Predicted vs observed mortality + fit GAM model

# prediction for gam
gam_pred <- predict(gam_fit)

# plot - observed mortality vs predicted mortality
#col <- c("Observed" == "#1abc9c", "Prediction" == "#e67e22")
ggplot(data, aes(x = data$Time)) + 
  geom_line(aes(y = data$Mortality, color = "Observed")) +
  xlab("Time") + ylab("Mortality") + ggtitle("Observed mortality vs predicted mortality") + 
  geom_line(aes(y = gam_pred ,color = "Prediction")) + 
  theme_bw()
```


GAM model:
```{r}
print(gam_fit)
summary(gam_fit)
gam_fit$sp
```


Spline component 

```{r, echo=FALSE}
plot(gam_fit)
```


## 1.4 Penalty function

```{r}

## 1.4 Penalty function

# The influance of penalty function
# increase lambda -> df_lambda decrease
# higher lambda higher penilize 
gam_smooth_low = smooth.spline(data$Time, data$Mortality, df = 25)
gam_smooth_high = smooth.spline(data$Time, data$Mortality, df = 100)

gam_smooth_low_pred = predict(gam_smooth_low)
gam_smooth_high_pred = predict(gam_smooth_high)
```


```{r, echo=FALSE}
ggplot() + 
  geom_line(aes(x = data$Time, y = data$Mortality, color = "Observed")) +
  xlab("Time") +
  ylab("Mortality") +
  ggtitle("Observed mortality vs predicted mortality") + 
  geom_line(aes(x = data$Time, y = gam_smooth_low_pred$y ,color = "Prediction low df")) +
  geom_line(aes(x = data$Time, y = gam_smooth_high_pred$y ,color = "Prediction high df"))

```


Degrees of freedom low:
```{r}
gam_smooth_low$lambda #2.147448e-05 
gam_smooth_low$df #25.00381
```

Degrees of freedom high:
```{r}
gam_smooth_high$lambda #2.579748e-08
gam_smooth_high$df #100.0082
```



Estimated deviance:
```{r}
gam_smooth_low 
gam_smooth_high
```


## 1.5 Plot residuals vs influenza
```{r, echo=FALSE}

## 1.5 Plot residuals vs influenza

col <- c("Observed" = "blue", "Residuals" = "red")
ggplot(data, aes(x = data$Time)) + 
  geom_line(aes(y = data$Mortality/5, color = "Observed")) +
  geom_line(aes(y = gam_fit$residuals, color = "Residuals")) +
  scale_y_continuous(sec.axis = sec_axis(~.*5, name = "Observed")) +
  xlab("Time") + ylab("Residuals") + ggtitle("Observed mortality and residuals") 

  #geom_line(aes(x = data$Time, y = gam_smooth_low_pred$y ,color = "Prediction low lambda")) +
  #geom_line(aes(x = data$Time, y = gam_smooth_high_pred$y ,color = "Prediction high lambda"))

```

## 1.6 Fit GAM model mortality is be modelled as an additive function
```{r}

# ## 1.6 Fit GAM model mortality is be modelled as an additive function

# gam_fit = gam(data$Mortality ~ data$Year +
#                 s(data$Week, k=length(unique(data$Week)),
#                   bs = "cp"), 
#               data = data,
#               method = "GCV.Cp")

# create new fitting model
gam_fit6 = gam(data$Mortality ~ data$Influenza +
                s(Year,k=length(unique(data$Year)), bs="gp")+
                s(Week,k=length(unique(data$Week)), bs="cp"),
              family = gaussian,
              data=data,
              method = "GCV.Cp")

# create prediction
gam_pred6 = predict(gam_fit6)

```

```{r, echo=FALSE}
# create visualization
col <- c("Observed" == "blue", "Prediction" == "green")
ggplot() + 
  geom_line(aes(x = data$Time, y = data$Mortality, color = "Observed")) +
  xlab("Time") +
  ylab("Mortality") +
  ggtitle("Observed mortality vs predicted mortality") + 
  geom_line(aes(x = data$Time, y = gam_pred6 ,color = "Prediction"))
```


```{r, warning=FALSE, message=FALSE, results='hide'}
#install.packages("pamr")
library(pamr)
#install.packages("kernlab")
library(kernlab)
#install.packages("glmnet")
library(glmnet)
#install.packages("kernlab")
library(kernlab)
#install.packages("sgof")
library(sgof)
```

# Assignment 2. High-dimensional methods
```{r, echo=FALSE, warning=FALSE, results='hide'}
# clean environment
rm(list = ls())
```

## 2.1 Nearest shrunken centriod classification

```{r, echo=FALSE}

## 2.1 Nearest shrunken centriod classification

# read data
# data = read.csv2("data.csv",
#                  fileEncoding = "ISO-8859-1")
data = read.csv2("data.csv")
```


```{r, echo=FALSE, results='hide'}
# devide data into train (70%) and test (30%) set - without scaling
n=dim(data)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.7)) 
train=data[id,]
test=data[-id,]
```

```{r}
#train
rownames(train) = 1:nrow(train)
x_train = t(train[,-4703]) # remove dependent variable
y_train = train[[4703]] # vector of the dependent variable
mydata_train = list(x = x_train,y=as.factor(y_train),geneid=as.character(1:nrow(x_train)), genenames=rownames(x_train))

#test 
rownames(test) = 1:nrow(test)
x_test = t(test[,-4703]) 
y_test = test[[4703]] 
mydata_test = list(x = x_test,y=as.factor(y_test),
                   geneid = as.character(1:nrow(x_test)), 
                   genenames = rownames(x_test))

# create the model
model = pamr.train(mydata_train,threshold=seq(0,4, 0.1))

# choice threshold by cv
cvmodel = pamr.cv(model,mydata_train)
```


Missclassification Error plot:

```{r, fig.height=9, echo=FALSE}
# settings {r, fig.height=9}
# Missclassification Error plot:
pamr.plotcv(cvmodel)
```


```{r, fig.height=9, echo=FALSE}
# Print out of the cvmodel:
print(cvmodel)
```



```{r}
# Min threshold:
# the threshold for the min error of cvmodel
best_thresbold = cvmodel$threshold[which.min(cvmodel$error)]
best_thresbold
```



```{r}
pamr.plotcen(model, mydata_train, threshold=best_thresbold)
```


The selected features - Threshold = 1.3: 


```{r, results='hide', message=FALSE}
# Number of features:
a = pamr.listgenes(model,mydata_train,threshold=best_thresbold)
```

```{r}
nrow(a)
```



```{r}
# List of the 10 most contributing features:

#cat(paste(colnames(data)[as.numeric(a[,1])], collapse='\n'))
top10 = as.matrix(colnames(data)[as.numeric(a[,1])][1:10])
top10
```

```{r}
# test error nearest shrunk
pred_model = pamr.predict(model,
                          newx = x_test,
                          threshold = 1) # also for the fit threshold of 1
cm_pred_model = table(y_test, pred_model)
test_error_nearestshrank = (cm_pred_model[1,2] + cm_pred_model[2,1]) / sum(cm_pred_model)
```

## 2.2 Test error & number of the contributing features

### 2.a Elasitc net

```{r, message=FALSE, warning=FALSE}

## 2.2 Test error & number of the contributing features

### 2.a Elasitc net

x_train = t(x_train) # transpose x_train back to normal
x_test = t(x_test)

# fit the elastic net 
elastic_net = cv.glmnet(x = x_train,
                        y = y_train,
                        family = "binomial",
                        alpha = 0.5)

# create prediction
elastic_net_pred = predict.cv.glmnet(elastic_net,
                                     newx = x_test,
                                     type = "class",
                                     s="lambda.min") 
# s penalty parameter 
cm_elastic_net  = table(y_test, elastic_net_pred)
test_error_elastic_net = (cm_elastic_net[1,2] + cm_elastic_net[2,1]) / sum(cm_elastic_net)

```

### 2.b Support vector machine

```{r, echo=FALSE}
# SVM 
# svm() function does not support vanilladot
# svm_fit = svm(x = x_train,
#               y = y_train,
#               kernel = "vanilladot")

# https://www.rdocumentation.org/packages/kernlab/versions/0.9-27/topics/ksvm
# used function - ksvm(), package
```


```{r, warning=FALSE, message=FALSE}
# SVM 
# used function - ksvm(), package
svm_fit  = ksvm(x = x_train,
                y = y_train,
                kernel = "vanilladot")
                #,scale = FALSE, # Variable(s) `' constant. Cannot scale data.
                #type = "C-svc") # C-svc C classification
                
svm_fit_pred = predict(svm_fit,
                       newdata = x_test)

cm_svm = table(y_test, svm_fit_pred)
test_error_svm = (cm_svm[1,2] + cm_svm[2,1]) / sum(cm_svm)
```

```{r, echo=FALSE}
# comparing the results 
# create df with the three values of the test error
test_error_df = data.frame(
  "Nearest shrank" = test_error_nearestshrank,
  "Elastic net" = test_error_elastic_net,
  "SVM" = test_error_svm
)


cf<-as.matrix(coef(elastic_net, elastic_net$lambda.min))
features_nearest_shrank = nrow(a)
features_svm_fit  = dim(data)[2] -1
features_elasticnet = length(names(cf[cf!=0,])) 

features_lengt = c(features_nearest_shrank,features_elasticnet, features_svm_fit)

test_error_df = rbind(test_error_df, features_lengt )

rownames(test_error_df)[1] = "Test error"
rownames(test_error_df)[2] = "Features"


# summary(model)
# svm_fit
# summary(elastic_net)
```


```{r, echo=FALSE}
knitr::kable(test_error_df, caption = "Comparative Table")
```


# 2.3 Benjamin-Hochberg method

```{r, message=FALSE, results='hide', echo=TRUE}
# Benjamin Hochberg
# laod the data

# data = read.csv2("data.csv",
#                  fileEncoding = "ISO-8859-1")

data = read.csv2("data.csv")
# create a data frame to save p-value & name
name_p_value = data.frame(name = character(),
                          p_value = numeric(),
                          stringsAsFactors = FALSE)
# for loop to fill data frame with wanted values
for (i in 1:4702) {
    x <- data[,i]
  p <- t.test( x ~ Conference,
               data = data,
               alternative = "two.sided" # default - don??t neet this
               )[["p.value"]]
  colname = colnames(data)[i]
 name_p_value = rbind(name_p_value, data.frame(colname,p))
}
colnames(name_p_value) = c("name", "p_value")

# -----
# TEST
# bla_test = BH(p_value, alpha = 0.05)
# bla_test_class = ifelse(bla_test$Adjusted.pvalues > 0.05, "Don??t reject", "Reject")
# which(bla_test_class == "Don??t reject")
# ----
# get the same results in the following calculation


# values to use the benjamin hochberg algorithm
alpha = 0.05
M = ncol(data)-1

# bring the p-values in order
# copy of the original data 
name_p_value_order = name_p_value
name_p_value_order = name_p_value_order[order(name_p_value_order$p_value),]

# calculate the prob for L
L <- c()
for (i in 1:M) {
  L[i] = name_p_value_order$p_value[i] - ((alpha *i)/ M)
}
# include L in the data frame name_p_value_order
name_p_value_order$L = L

# save rejecteion region
rejection_switch_vector <- which(name_p_value_order$L < 0)
rejection_switch_max <- max(which(L < 0))

reject_yes_no <- c()
for (i in 1:length(L)) {
  if( i <= rejection_switch_max){
    reject_yes_no[i] = "yes"
  } else{
    reject_yes_no[i] = "no"
  }
}

# add vector rejection_yes_no to data frame name_p_value_order
name_p_value_order$rejected = reject_yes_no

# create data frame just with rejected variables 
name_p_value_order_rejected = name_p_value_order[1:rejection_switch_max,]
```


List of rejected features: 
```{r, echo=FALSE}
name_p_value_order_rejected
```

In this table are we able to see the 39 features which got rejected. This means, all the features have a clear relation to the conference text. 
As well do we see the p-value, the calculated L-value (Benjamini Hochberg) and if the feature to rejected or not.

\newpage
```{r ref.label=knitr::all_labels(), echo = T, eval = F}
```

